The ``lifecycle methods'' are the special named methods of class based components that get called by React. We've already been using \texttt{constructor} and \texttt{render}, but there are some other useful ones to cover.


\section{\texttt{constructor}}

The \texttt{constructor} method is not a React-specific method, it's part of JavaScript, but it does have a very specific use for React components.
\\

It should only be used for:

\begin{itemize}
    \item Setting up the initial state
    \item Binding methods
\end{itemize}

If you don't need to do either of these then you don't need to write a \texttt{constructor} method at all and it will fallback to the parent's (\texttt{Component}) \texttt{constructor}.
\\

Doing anything else inside the \texttt{constructor} could potentially delay rendering, which will lead to a poor user experience. We'll look at a better approach in a minute.


\section{\texttt{render}}

The \texttt{render} method is called by React when it needs to re-render the component – when either \texttt{state} or \texttt{props} change. It should return valid JSX. The only code that should go in \texttt{render} should be related to generating the JSX. It should not make changes to the DOM or have any other side-effects, as these can slow down rendering and potentially create all sorts of bugs.



\section{\texttt{componentDidMount}}

This method is called for us by React just after the component first renders, so any code we put in it will run when a component is first created.\footnote{If a component is removed and then re-added, the \texttt{componentDidMount} method will run again.} This is \href{https://medium.com/@mahcloud/actions-in-the-constructor-or-componentdidmount-be3720e4a9a6}{subtly different} from the \texttt{constructor} as it doesn't delay \texttt{render}.
\\

\texttt{componentDidMount} is a safe place to perform DOM manipulation and other side-effects.
\\

We could use this, for example, to setup a timer so that our component hides itself after 5 seconds:

\begin{minted}{javascript}
    componentDidMount() {
        this.timerID = setTimeout(() => {
            // assumes logic in JSX to hide component
            this.setState({ hide: true });
        }, 5000);
    }
\end{minted}

Sometimes it's also necessary to add event handlers to DOM elements that aren't part of the component, like the \texttt{window} object. We can set this up in \texttt{componentDidMount} too:

\begin{minted}{javascript}
    componentDidMount() {
        // a drag-and-drop component might need to know when the mouse
        // is no longer held down, no matter what component it's over
        window.addEventListener("mouseup", this.handleStopDragging);
    }
\end{minted}

If you do setup timers or event handlers in \texttt{componentDidMount}, make sure you remove them using \texttt{componentWillUnmount}.


\section{\texttt{componentWillUnmount}}

React does all sorts of clever stuff for us, including adding and removing all necessary event listeners to DOM elements that it creates. However, if we set something up in \texttt{componentDidMount}, such as adding event listeners, it's our responsibility to remove them again when the component is no longer needed.

For example, clearing a timer so it doesn't keep running:

\begin{minted}{javascript}
    componentWillUnmount() {
        // make sure the timer doesn't keep running
        // using the timerID property setup in componentDidMount
        clearTimer(this.timerID);
    }
\end{minted}

Or removing event listeners:

\begin{minted}{javascript}
    componentWillUnmount() {
        // we need to remove any event listeners
        // that React hasn't setup itself
        window.removeEventListener("mouseup", this.handleStopDragging);
    }
\end{minted}

If we forget to do this, over the life of our app multiple copies of the same timer or event listener will build up, using memory and CPU unnecessarily (and potentially leading to all sorts of weird bugs).


\section{\texttt{componentDidUpdate}}

Sometimes we want to run a bit of code every time the component updates (either props or state), but \textit{after} it has finished rendering.
\\

Again, \texttt{componentDidUpdate} is a safe place to perform things that might take a while, perform DOM manipulation, or create other side-effects.
\\

For example, say we wanted to autosave the state in \href{https://developer.mozilla.org/en-US/docs/Web/API/Window/localStorage}{local storage} whenever \textit{any} value in state changed:

\begin{minted}{javascript}
    componentDidUpdate() {
        // store the changes in localStorage whenever any value in state changes
        // could load them again in componentDidMount
        localStorage.setItem('myComponent', JSON.stringify(this.state));
    }
\end{minted}

Make sure you don't use \texttt{setState} inside \texttt{componentDidUpdate} unless it's inside a conditional, otherwise you'll create an infinite loop.

\hr

If you're  using any of the lifecycle methods to perform DOM manipulation, \textbf{make sure there isn't a React way to achieve the same thing first}. As a general rule of thumb, if the DOM manipulation is on DOM elements that are part of the component (or its children), you should be able to use React-specific methods.



\section{Additional Resources}

\begin{itemize}[leftmargin=*]
    \item \href{https://reactjs.org/docs/react-component.html#the-component-lifecycle}{React: The Component Lifecycle}
    \item \href{https://reactjs.org/docs/react-component.html#constructor}{React: \texttt{constructor}}
    \item \href{https://reactjs.org/docs/react-component.html#render}{React: \texttt{render}}
    \item \href{https://reactjs.org/docs/react-component.html#componentdidmount}{React: \texttt{componentDidMount}}
    \item \href{https://reactjs.org/docs/react-component.html#componentwillunmount}{React: \texttt{componentWillUnmount}}
    \item \href{https://reactjs.org/docs/react-component.html#componentdidupdate}{React: \texttt{componentDidUpdate}}
\end{itemize}
