If a React app needs to work with data that's stored on a server somewhere, it will need to make API calls.
\\

So far the only way we know to get data in the browser is to navigate to a new page and the only way to submit data is to create a form with a \texttt{POST} method. Both of these involve refreshing the page, which is no good for an app that runs purely in JavaScript.
\\


``AJAX'' (\textbf{Asynchronous JavaScript and XML}) is an outdated term for doing HTTP requests using JavaScript.
\\

The name comes from the pre-JSON era, when XML was the most common way to transfer data:\footnote{XML is still used in some legacy systems}

\begin{minted}{xml}
    <employees>
        <employee>
            <name>Jenny</name>
            <email>jenny1987@gmail.com</email>
        </employee>
        <employee>
            <name>Rahul</name>
            <email>rahul12@gmail.com</email>
        </employee>
        <employee>
            <name>Mia</name>
            <email>mia87@gmail.com</email>
        </employee>
    </employees>
\end{minted}

As you can see, XML is verbose and repetitive. Nowadays JSON is used pretty much everywhere. So it should really be called ``AJAJ'', but that doesn't sound as cool.
\\

In any case, people still say ``AJAX'' when they're talking about JavaScript making HTTP requests.


\section{Axios}

There are various ways to make an AJAX request from the browser.
\\

The original method, \texttt{XMLHttpRequest}, is horrifying to use, as it pre-dates the time when people realised that JavaScript didn't have to be unpleasant to work with.
\\

The more modern \texttt{fetch} API is much nicer to work with and is \href{https://caniuse.com/#feat=fetch}{now supported in most browsers}. However, it's still a bit ungainly to work with.
\\

We'll be using \href{https://github.com/mzabriskie/axios}{Axios} for our HTTP requests. It's a handy library that makes doing HTTP requests with JavaScript really easy.\footnote{And, as an added bonus, it works in Node too.}
\\

You can install it with:

\begin{minted}{js}
    npm install axios
\end{minted}


\begin{infobox}{Axios 0.19}
    Axios 0.19 introduced a bug which means that the \texttt{params} property of \\ \texttt{axios.create} gets lost. This bug was reported on GitHub six months ago, but remains unfixed.
    \\

    \textbf{For now you will want to run \texttt{npm install axios@0.18} to force NPM to install an older, but working, version of Axios}.
\end{infobox}


Axios provides us with methods for the various HTTP methods:

\begin{minted}{js}
    import axios from "axios";

    // make a GET request
    axios.get("/articles");

    // make a POST request, with the given data
    axios.post("/articles", {
        title: "Hello",
        article: "Blah blah blah",
        tags: ["fish", "cat", "spoon"],
    });

    // make a PUT request, with the given data
    axios.put("/articles/5", {
        title: "Hello Again",
        article: "Blah blah blah!",
        tags: ["fish", "cat"],
    });

    // make a DELETE request
    axios.delete("/articles/4");
\end{minted}


\subsection{Config}

One of Axios' best features is it's easy to configure the parts of your HTTP request that are the same every time.
\\

For example, if we're using a RESTful API, our base URL, API key, and the \texttt{Accept} header are going to be the same for every request, so we can setup a version of Axios that has those already setup:

\begin{minted}{js}
    // import the library version of axios
    import axios from "axios";

    // create a version of axios with useful defaults
    export default axios.create({
        // use your own url
        baseURL: "http://username.restful.training/api/",

        // use your own key
        params: {"key": "4fdea06a65ba1491091c0db709faf0cce944067a"},

        // make sure we get JSON back
        headers: {"Accept": "application/json"},
    });
\end{minted}

Then we can import \textit{that} file and use it as follows:

\begin{minted}{js}
    // import *local* version of axios
    import axios from "./axios";

    // automatically handle base URL and key
    axios.get("/articles");
\end{minted}



\section{Additional Resources}

\begin{itemize}[leftmargin=*]
    \item \href{https://github.com/axios/axios}{Axios}
    \item \href{https://blog.logrocket.com/how-to-make-http-requests-like-a-pro-with-axios/}{How To Make HTTP Requests Like a Pro with Axios}
\end{itemize}
