When we make a request to the API we don't get a response straight away. Even on a fast connection it can take 100ms or more to get a response. That might not sound like a long time, but for a computer that's ages.
\\

So, JavaScript doesn't just stop doing everything until you get a response back: it keeps on doing whatever else needs to be done until it gets a response.
\\

When we don't get back a response immediately, we need to deal with the response \textbf{asynchronously}: in other words, the code that runs when the response comes back could run at \textit{any time} in the future (or never).  We've actually dealt with a similar concept when we wrote event handlers.


\section{Promises}

So, when we make a request with Axios, we can't be given back the response to store in a variable, as the response doesn't exist yet. What we get back instead is a \textbf{Promise}.
\\

A Promise is an object with a \texttt{.then()} method. We can pass the \texttt{.then()} method a function, which will run once the response comes back. With Axios, the function will be passed the response as the first parameter.

\begin{minted}{js}
    // make the request
    let promise = axios.get("/articles");

    // setup the handler for when the response is successful
    promise.then(response => {
        console.log(response);
    });
\end{minted}

Rather than using an intermediary variable, generally we use chaining with Promises:

\begin{minted}{js}
    // use destructuring to get the data property
    axios.get("/articles").then(({ data }) => {
        console.log(data);
    });
\end{minted}

Be aware that \textbf{there's no guarantee the promises will resolve in the order you wrote them}. For example, \texttt{GET} requests tend to be quicker than other types of request as they generally involve less server activity.


\subsection{Errors}

Sometimes things will go wrong: you might have sent invalid data, the API server might be down, or any number of other issues.
\\

The \texttt{.then()} method of a Promise actually accepts a second argument which will be called if something goes wrong.

\begin{minted}{js}
    axios.get("/articles").then(response => {
        console.log("Everything has worked");
    }, error => {
        console.log(error); // logs an error message
        console.log(error.response); // the response object
        console.log("Something has gone wrong");
    });
\end{minted}

Alternatively you can use the \texttt{.catch()} method:

\begin{minted}{js}
    axios.get("/articles").then(response => {
        console.log("Everything has worked");
    }).catch(error => {
        console.log("Something has gone wrong", error.response);
    });
\end{minted}

You could use the above methods to handle form validation errors and the like.

\section{Additional Resources}

\begin{itemize}[leftmargin=*]
    \item \href{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Global_Objects/Promise}{MDN: Promises}
    \item \href{http://exploringjs.com/es6/ch_promises.html}{Promises: A Technical Look}
    \item \href{https://developer.mozilla.org/en-US/docs/Web/JavaScript/Reference/Statements/async_function}{MDN: \texttt{async} functions}
    \item \href{https://www.smashingmagazine.com/2019/10/asynchronous-tasks-modern-javascript/}{Writing Asynchronous Tasks In Modern JavaScript}
\end{itemize}
